---
title: "Using the `Powers` package"
author: "Cassandra Konecny"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the `Powers` package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

__Note__: This package isn't actually meant to be "serious". I made it just for teaching purposes.

Want to compute powers without having to use `^` or `**`? Give `sqrt` some friends by loading the `powers` package! 

```{r}
library(Powers)
```

Want squares, cubes, or reciprocals? Easy.

```{r}
square(1:10)
cube(1:10)
reciprocal(1:10)
```

`apply`-family friendly, too!

```{r}
my_list <- list(1:10, 0.5, -0.7)
## So base-R-boring!
lapply(my_list, function(x) x^2)
## Use powers instead!
lapply(my_list, square)
```

It even comes with the ability to display its output in plot form!

```{r}
reciprocal(1:10, plot_it=TRUE)
```

Wow!!!

AND when powers don't make you feel Powerful enough there is the regress function! This let's you perform a linear regression between two variables and outputs a plot of the results.

```{r}
## Let's say that we have one vector of all of our favourite numbers
fav <- c(1,5,7,9,19)
## And one of all of our least favourite numbers
unfav <- c(3,6,8,2,4)
## we can use regress() to plot fav as a function of unfav
regress(fav, unfav, TRUE)
## The lack of relationship is overwhelming!
```
But what if we wanted the difference between the numbers instead? Fear not - diff() is here!

```{r}
## Let's calculate the difference between vectors!
diff(fav,unfav)
```


