---
title: "hw_02_gapminder_dplyr"
author: "Cassandra"
date: '2017-09-26'
output:
  word_document: default
  pdf_document: default
  html_document: default
---

# Bring rectangular data in
## Install and load required packages
```{r}
install.packages("gapminder")
library(gapminder)
install.packages("tidyverse")
library(tidyverse)
```


# Smell test the data
## Is it a data.frame, a matrix, a vector, a list?
A data.frame is a list of vectors
```{r}
data <- gapminder
typeof(data)
```
## What’s its class?
The class of the data.frame is 
```{r}
class(data)
```
## How many variables/columns?
There are 6 variables/columns in the gapminder data set.
```{r}
ncol(data)
```

## How many rows/observations?
There are 1704 rows/observations in the gapminder data set.

```{r}
nrow(data)
```
## Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?
Get dimmensions of the data set using dim(). This would be good if we want a really quick overview of the size of the dataset
```{r}
dim(data)
```
Get length of object - in this case the length of the list of vectors (rows). This would be useful if we only wanted to know the number of rows (similar function to the nrow() function)
```{r}

length(data)
```
Get the names of the columns - we can see that there are 6 in total. This is useful if we not only want the number of columns, but also their names.
```{r}
names(data)
```
## What data type is each variable?
Use the typeof() function to find the data type of each variable. We can use the column names given above

country:
```{r}
typeof(data$country)
```
continent:
```{r}
typeof(data$continent)
```
year:
```{r}
typeof(data$year)
```
lifeExp:
```{r}
typeof(data$lifeExp)
```
pop:
```{r}
typeof(data$pop)
```
gdpPercap:
```{r}
typeof(data$gdpPercap)
```


